Bioconductor workflow for single-cell RNA-seq data analysis: dimensionality reduction, clustering, and pseudotime ordering

### 2017-06-19 workflow_combineMinSize10_mds/workflow.Rmd
combineMinSize=10 and mds

```{r rsec_50}
fn <- '../data/RSEC_W.rda'
if (runClus & !file.exists(fn)){
  #symbol for samples missing from original clustering
  seObj <- SummarizedExperiment(t(W), colData = colData(core))
  print(system.time(ceObj <- RSEC(seObj, k0s = 4:15, alphas = c(0.1),
                                  betas = 0.8, dimReduce="none",
                clusterFunction = "hierarchical01", minSizes=1,
                ncores = NCORES, isCount=FALSE,
                subsampleArgs = list(resamp.num=100,
                                     clusterFunction="kmeans",
                                     clusterArgs=list(nstart=10)),
                seqArgs = list(k.min=3, top.can=5), verbose=TRUE,
                combineProportion = 0.7,
                mergeMethod = "none", random.seed=424242)))
  save(ceObj, file = fn)
}else{
  load(fn)
}
```

@@

Cluster name | Description | Color | Correspondence
-------------|-------------|-------| ----------
c1 | HBC | blue | original 1, 5
c2 | GBC | immature neurons | MV 1 | green | original 2, 3, 11
c3 | Sus | red | original 4, 7
c4 | Neuron | orange | original 9, 12
c5 | Immature Neuron | purple | original 10, 14
c6 | new and small | brown | original 9
c7 | Microvillus | cyan | original 15

@@

### Use previous W

@@

#### Unsupervised

@@

K = 4 two lineages: sus is right after HBC, differ only in end point (neuron, new cluster)

K = 5 three lineages: hbc-gbc-immature neuron-new-neuron; hbc-gbc-mv; hbc-sus (perfect!)

@@

#### Supervised

K = 3 four lineages: might be good, three lineages are correct and the fourth is plausible.

K = 5 same as K=3

@@

  lineages <- get_lineages(X, clus.labels = cl, start.clus = "c1",
                           end.clus = c("c3", "c6", "c7"))
  curves <- get_curves(X, clus.labels = cl, lineages = lineages)
  plot_curves(X, cl, curves, col.clus = pal)
  plot_tree(X, cl, lineages, col.clus = pal)

@@

### Re-fitting zinbwave

#### Unsupervised

K = 3 four lineages: immature neurons as a end point (not good)

K = 4 same as K = 3

K = 5 same as K = 3

```{r refit_zinb}
fn <- '../data/refit_zinbwave_slingshot.rda'

@@

#### Supervised

K = 3 four lineages (the small new cluster is still an end point)

K = 5 three lineages (correct)

@@

  lineages <- get_lineages(X, clus.labels = cl, start.clus = "c1",
                           end.clus = c("c3", "c6", "c10"))
  curves <- get_curves(X, clus.labels = cl, lineages = lineages)
  plot_curves(X, cl, curves, col.clus = pal)
  plot_tree(X, cl, lineages, col.clus = pal)

@@

CONCLUSION: K = 5 is never great as GBC is generally an end cluster. K = 4 is ok for all the methods and a bit better when zinbwave is refitted. K = 3 when refitting and supervized is good.

It seems to me that using slingshot on W without re-fitting zinbwave with k = 4 gives good results where supervized mode is slightly better than unsupervized. It is just a one shot example and we should obviously not make a general conclusion, but I think that for the purpose of the workflow it is fine to use slingshot without refitting zinbwave. We should write a note to the user that it is better to refit zinbwave to have more power.

>>>>

```{r rsec_50}
fn <- '../data/RSEC_W_combineMinSize10.rda'
if (runClus & !file.exists(fn)){
  #symbol for samples missing from original clustering
  seObj <- SummarizedExperiment(t(W), colData = colData(core))
  print(system.time(ceObj <- RSEC(seObj, k0s = 4:15, alphas = c(0.1),
                                  betas = 0.8, dimReduce="none",
                clusterFunction = "hierarchical01", minSizes=1,
                ncores = NCORES, isCount=FALSE,
                subsampleArgs = list(resamp.num=100,
                                     clusterFunction="kmeans",
                                     clusterArgs=list(nstart=10)),
                seqArgs = list(k.min=3, top.can=5), verbose=TRUE,
                combineProportion = 0.7,
                mergeMethod = "none", random.seed=424242,
                combineMinSize = 10)))
  save(ceObj, file = fn)
}else{
  load(fn)
}
```

@@

Cluster name | Description | Color | Correspondence
-------------|-------------|-------| ----------
c1 | HBC | blue | original 1, 5
c2 | GBC | immature neurons | MV 1 | green | original 2, 3, 11
c3 | Sus | red | original 4, 7
c4 | Neuron | orange | original 9, 12
c5 | Immature Neuron | purple | original 10, 14
c6 | Microvillus | cyan | original 15

@@

### Use W components
#### Use previous W

@@

##### Unsupervised

@@

K = 5 three lineages: hbc-gbc-immature-neuron; hbc-gbc-mv; hbc-sus (perfect!)

@@

##### Supervised

K = 3, 4, 5 get same results, all good.

@@

  lineages <- get_lineages(X, clus.labels = cl, start.clus = "c1",
                           end.clus = c("c3", "c6"))
  curves <- get_curves(X, clus.labels = cl, lineages = lineages)
  plot_curves(X, cl, curves, col.clus = pal)
  plot_tree(X, cl, lineages, col.clus = pal)

@@

#### Re-fitting zinbwave

##### Unsupervised

K = 3, 4, 5 get same results, all good.

```{r refit_zinb}
fn <- '../data/refit_zinbwave_slingshot_combineMinSize10.rda'

@@

##### Supervised

K = 3, 4, 5 get same results, all good.

@@

  lineages <- get_lineages(X, clus.labels = cl, start.clus = "c1",
                           end.clus = c("c3", "c6"))
  curves <- get_curves(X, clus.labels = cl, lineages = lineages)
  plot_curves(X, cl, curves, col.clus = pal)
  plot_tree(X, cl, lineages, col.clus = pal)

@@

### Use MDS of W

Here, instead of using the first components of W, we use the MDS of W.

#### Use previous W

The input of slingshot is the MDS of W used for clusterExperiment where the number of dimensions is k where k in (3, 4, 5) here.

##### Unsupervised

K = 4, 5 now, it is ok, three lineages: hbc-gbc-immature-neuron; hbc-gbc-mv; hbc-sus (perfect!)

```{r}
our_cl <- primaryClusterNamed(ceObj)
cl = our_cl[our_cl != "-1"]
pal = pal[names(pal) != '-1']
```

```{r}
for (k in Kvec){
  X <- W[our_cl != "-1", ]
  mds <- cmdscale(dist(X), eig = TRUE, k = k)
  X <- mds$points

##### Supervised

K = 3, 4, 5 get same results, all good.

```{r}
for (k in Kvec){
  X <- W[our_cl != "-1", ]
  mds <- cmdscale(dist(X), eig = TRUE, k = k)
  X <- mds$points

  lineages <- get_lineages(X, clus.labels = cl, start.clus = "c1",
                           end.clus = c("c3", "c6"))
  curves <- get_curves(X, clus.labels = cl, lineages = lineages)
  plot_curves(X, cl, curves, col.clus = pal)
  plot_tree(X, cl, lineages, col.clus = pal)

#### Re-fitting zinbwave

##### Unsupervised

K = 3, 4, 5 get same results, all good.

```{r}
for(k in Kvec) {
  X <- getW(zinbList[[k - 2]])
  mds <- cmdscale(dist(X), eig = TRUE, k = k)
  X <- mds$points

##### Supervised

K = 3, 4, 5 get same results, all good.

```{r}
for(k in Kvec){
  X <- getW(zinbList[[k - 2]])
  mds <- cmdscale(dist(X), eig = TRUE, k = k)
  X <- mds$points

  lineages <- get_lineages(X, clus.labels = cl, start.clus = "c1",
                           end.clus = c("c3", "c6"))
  curves <- get_curves(X, clus.labels = cl, lineages = lineages)
  plot_curves(X, cl, curves, col.clus = pal)
  plot_tree(X, cl, lineages, col.clus = pal)
